A Formal Model of Checked C

### 2022-07-17 sandp23/background.tex

\begin{figure}[t]
{\small
  \begin{lstlisting}[xleftmargin=4 mm]
//in checked region, tainted version
int tainted_compare_1(
  nt_array_tptr<char> x : count (0),
  nt_array_tptr<char> y : count (0)) {
  int len_x = strlen(x);
  int len_y = strlen(y);
  nt_array_ptr<char> tx : count (len_x)
   = malloc(nt_array<char>, len_x);
  nt_array_ptr<char> ty : count (len_y)
   = malloc(nt_array<char>, len_y);
  safe_memcpy(tx,x,len_x);
  safe_memcpy(ty,y,len_y);
  return compare_1(tx,ty);
}
...

//calling the function turns
//an unchecked region to a checked region.
int tainted_stringsort(nt_array_tptr
     <nt_array_tptr<char>> s : count (n),
 tptr<(int)(nt_array_tptr<char>,
  nt_array_tptr<char>)> cmp, int n) {
  int i;
  nt_array_ptr<nt_array_ptr<char>> p : count (n)
    = malloc(nt_array<nt_array_ptr<char>>, n);
  for(i = 0; i < n; i++) {
    int len = strlen s[i];
    nt_array_ptr<char> tmp : count (len)
      = new malloc(nt_array<char>, len);
    safe_memcpy(tmp,s[i],len);
    p[i] = tmp;
  }
  ptr<(int)(nt_array_ptr<char> : count (0),
  nt_array_ptr<char> : count (0))>
    cfun = find_check(cmp);

  return stringsort(p,cfun);
}
  \end{lstlisting}
}
\caption{Tainted stringsort Code}
\label{fig:checkedc-example-2}
\end{figure}

@@

Conceptually, the shell is run in a checked region.
One can imagine that a tainted shell is a safe closure that contains a checked block.
Once the closure is called, the system is turned to a checked region,
so that the checked function call at \Cref{fig:checkedc-example-2} line 13 is safe, 
even if it is called by $g$ in \Cref{fig:checkedc-example-3} , because it lives in the checked region.
In the \systemname formalism, we formalize a \textit{checked} block on top of the existing unchecked regions,
and the transition of a tainted shell call creates a checked block containing the shell body.
We also make sure that no arguments in these tainted shell contains any checked pointers,
as well as no output is of a checked type.

>>>>

\begin{figure}[t]
{\small
  \begin{lstlisting}[xleftmargin=4 mm]
//in checked region, tainted version
int tainted_compare_1(
  nt_array_tptr<char> x : count (0),
  nt_array_tptr<char> y : count (0)) {
 checked (x,y) {
  int len_x = strlen(x);
  int len_y = strlen(y);
  nt_array_ptr<char> tx : count (len_x)
   = malloc(nt_array<char>, len_x);
  nt_array_ptr<char> ty : count (len_y)
   = malloc(nt_array<char>, len_y);
  safe_memcpy(tx,x,len_x);
  safe_memcpy(ty,y,len_y);
  return compare_1(tx,ty);
 }
}
...

//calling the function turns
//an unchecked region to a checked region.
int tainted_stringsort(nt_array_tptr
     <nt_array_tptr<char>> s : count (n),
 tptr<(int)(nt_array_tptr<char>,
  nt_array_tptr<char>)> cmp, int n) {
 checked (s,cmp,n) {
  int i;
  nt_array_ptr<nt_array_ptr<char>> p : count (n)
    = malloc(nt_array<nt_array_ptr<char>>, n);
  for(i = 0; i < n; i++) {
    int len = strlen s[i];
    nt_array_ptr<char> tmp : count (len)
      = new malloc(nt_array<char>, len);
    safe_memcpy(tmp,s[i],len);
    p[i] = tmp;
  }
  ptr<(int)(nt_array_ptr<char> : count (0),
  nt_array_ptr<char> : count (0))>
    cfun = find_check(cmp);

  return stringsort(p,cfun);
 }
}
  \end{lstlisting}
}
\caption{Tainted stringsort Code}
\label{fig:checkedc-example-2}
\end{figure}

@@

Conceptually, the shell is run in a checked region.
One can imagine that a tainted shell is a safe closure that contains a checked block.
Once the closure is called, the system is turned to a checked region,
so that the checked function call at \Cref{fig:checkedc-example-2} line 14 is safe, 
even if it is called by $g$ in \Cref{fig:checkedc-example-3} , because it lives in the checked region.
In the \systemname formalism, we formalize a \textit{checked} block on top of the existing unchecked regions,
and the transition of a tainted shell call creates a checked block containing the shell body.
We also make sure that no arguments in these tainted shell contains any checked pointers,
as well as no output is of a checked type.
